Parsing Postfix log files

### 2009-01-16 doc/logparser-results.tex
Improvements after meeting Carl.
Converted from article to report, and sections to chapters. Some cleanups. @ doesn't need escaping in glossary entries with new glossaries package.
Further corrections and improvements from proofreading.
Lots of minor improvements and cleanups to the State of the Art Review,     including running diction on it. Checked that I've written 'log files' and 'log lines' everywhere, instead of     'files' or 'lines', and added a make target to help with it.  This is     supported by paragraphs-to-lines, which will hopefully find other uses.  I     renamed a couple of

\section{Results}

@@

@@

\subsection{Parser Efficiency}

@@

\subsubsection{Architecture Scalability: Input Size}

An important property of a parser is how parsing time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
\Graphref{parsing time vs file size vs number of lines graph} shows the
parsing time in seconds, file size in MB, and tens of thousands of log
lines for each of the \numberOFlogFILES{} log files.  All three lines run
roughly in parallel, giving the impression that the algorithm scales
linearly with input size.  This impression is borne out by
\graphref{parsing time vs file size vs number lines factor} which plots
both the ratio of file size vs parsing time, and the ratio of number of log
lines vs parsing time (higher is better in both cases); \tableref{parsing
time vs file size vs number lines factor table} shows the same ratios for
different groups of log files.  The ratios are quite tightly banded,
providing empirical evidence that the algorithm scales linearly.  The ratio
increases (i.e.\ improves) for log files 22 and 62--68 despite their larger
than average size (shown in \graphref{parsing time vs file size vs number
of lines graph}); this is explained fully in \sectionref{Explaining the
peaks in log file size}.

\showtable{build/include-file-size-and-number-of-lines-vs-parsing-time}{Ratio
of file size and number of lines to parsing time}{parsing time vs file size
vs number lines factor table}

\showgraph{build/graph-input-size-vs-parsing-time}{Parsing time, file size,
and number of lines}{parsing time vs file size vs number of lines graph}

\showgraph{build/graph-input-size-vs-parsing-time-ratio}{Ratio of file size
and number of lines to parsing time}{parsing time vs file size vs number
lines factor}

\subsubsection{Explaining the peaks in log file size}

@@

\Graphref{parsing time vs file size vs number of lines graph} shows log
file size varying between 30--100MB, except for log files 22 (200MB) and
62--68 (400--900MB).  These large log files were caused by a user
unintentionally creating a mail forwarding loop, resulting in a huge number
of mail deliveries and up to an order of magnitude more log lines than is
usual.  The user specified his mail forwarding as: \newline{}
\tab{}\texttt{$\backslash$username, username@domain} \newline{} This
instructs Postfix to deliver the mail to the local user, and also forward a
copy to the remote address; this is generally not a problem except that the
remote address in this instance is the address the mail was originally sent
to, creating an infinite loop.  To prevent this happening Postfix examines
the Delivered-To header in the mail, and if the mail has already been
delivered to the current address it is bounced back to the sender with the
error message \newline{} \tab{} \texttt{mail forwarding loop for
username@domain}\newline{}  Ordinarily this works well, but unfortunately
in this instance the user noticed they had not received any mail in a while
and opted to send a test mail to themselves, causing a loop not caught by
Postfix:

XXX REWRITE THIS SECTION SO THAT IT'S CORRECT\@: POINT 4 SAYS THAT THE
ENVELOPE SENDER CHANGES, IS THAT CORRECT\@?

@@

\subsubsection{Rule ordering for efficiency}

@@

Differences in rule ordering have less effect on parsing time when parsing
log files 22 and 62--68, due to the different distribution of log lines in
those log files (see \sectionref{Explaining the peaks in log file size}).
When a mail loop occurs the vast majority of log lines are from Postfix
components that have few rules associated with them, whereas in general the
most log lines are from Postfix components that have a high number of
associated rules; this results in the average number of rules required to
parse a log line from log files 22 or 62--68 being considerably lower than
usual.  The ratio of number of lines versus parsing time in
\graphref{parsing time vs file size vs number lines factor} shows a
noticeable improvement at log files 22 and 62--68, where a higher ratio
means more log lines parsed per unit of time; similarly \graphref{Parsing
time relative to shuffled ordering graph} shows a much smaller difference
between the orderings at log files 22 and 62--68.

\subsubsection{Scalability as the number of rules rises}

@@

\subsubsection{Caching compiled regexs}

@@

Two large dips can be seen in \graphref{Increase in parsing time when not
caching compiled regexes graph} at log files 22 and 62--68, corresponding
to the spikes in log file size in \graphref{parsing time vs file size vs
number of lines graph}.  The distribution of log lines across rules when
there is a mail loop is much different to the norm, and the average number
of rules consulted per log line is much lower: this results in far fewer
regex compilations per line when there is a mail loop, and a
corresponding decrease in the average parsing time for a single log line.
The outcome is that caching compiled regexes is proportionally less
important when the log file contents were created by a mail loop.  The
increases in parsing time when not caching compiled regexes for
different groups of log files are summarised in \tableref{Increase in
parsing time when not caching compiled regexes table}.

@@

\subsubsection{Summary}

@@

\subsection{Coverage}

@@

\subsubsection{Log lines covered}

@@

Full coverage of log lines is easy to achieve yet hard to maintain.  It is
easy to achieve full coverage for a limited set of log files (at the time
of writing the parser is tested with \numberOFrules{} rules, fully parsing
\numberOFlogFILESall{} contiguous log files from Postfix 2.2 through to
Postfix 2.5), and new rules are easy to add.  Maintaining full coverage is
hard because other servers have different restrictions with custom
messages, \gls{DNSBL} messages change over time, major releases of Postfix
change warning messages (usually adding more information), etc.,\ so over
time the log lines drift and change.  \Graphref{rule hits graph} shows the
number of hits for each rule over all \numberOFlogFILES{} log files.  The
number of hits per rule is quite unevenly spread, resembling a Power Law
distribution --- it is obvious that a small number of rules match the vast
majority of the lines, and more than half the rules match fewer than 100
times.

@@

\subsubsection{Mails covered}

@@

Whereas verifying by inspection that the parser correctly deals with all
60,721,709 lines in the test log files is infeasible, verifying a subset of
those log files is a tractable, if extremely time consuming, task.  A
sample of log lines was obtained by randomly selecting a log file:

@@

The first 6000 lines of this log file (roughly 0.01\% of the total number
of log lines used in testing) was extracted:

@@

It is important that the log lines used are contiguous so that all log
entries are present for as many of the connections as possible.  This log
segment was parsed with all debugging options enabled, resulting in 167,448
lines of output.\footnote{A mean of 27.908 lines of output per line of
input; each connection has 30 debugging lines, plus 21 debugging lines per
result.  Connections which have been cloned will have the cloned connection
in their debugging output, plus another 33 debugging lines.  Those numbers
are approximate, and may vary $\pm{}$ 2.  There is a linear relationship
between the number of log lines and debugging lines: $33(connections) +
30(accepted~~mails) + 21(results)$.  This formula is an approximation only,
and has not been rigorously verified.}  All 167,448 lines were examined in
conjunction with the input log file and a dump of the resulting database,
verifying that for each of the input lines the parser used the correct rule
and executed the correct action, that in turn produced the correct result
and inserted the correct data in the database.  The log file segment
produced 4 warnings, 10 mails remaining in the state tables, and 1625
connections correctly entered in the database.

@@

>>>>

\chapter{Results}

@@

XXX COMPARE BEST AND WORST POSSIBLE RESULTS WITH REAL RESULTS\@.

@@

\section{Parser Efficiency}

@@

\clearpage{}

\subsection{Architecture Scalability: Input Size}

An important property of a parser is how parsing time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
\Graphref{parsing time vs file size vs number of log lines graph} shows the
parsing time in seconds, file size in MB, and tens of thousands of log
lines for each of the \numberOFlogFILES{} log files.  All three lines run
roughly in parallel, giving the impression that the algorithm scales
linearly with input size.  This impression is borne out by
\graphref{parsing time vs file size vs number of log lines factor} which
plots both the ratio of file size vs parsing time, and the ratio of number
of log lines vs parsing time (higher is better in both cases);
\tableref{parsing time vs file size vs number of log lines factor table}
shows the same ratios for different groups of log files.  The ratios are
quite tightly banded, providing empirical evidence that the algorithm
scales linearly.  The ratio increases (i.e.\ improves) for log files 22 and
62--68 despite their larger than average size (shown in \graphref{parsing
time vs file size vs number of log lines graph}); this is explained fully
in \sectionref{Explaining the peaks in log file size}.

\showtable{build/include-file-size-and-number-of-log-lines-vs-parsing-time}{Ratio
of file size and number of log lines to parsing time}{parsing time vs file
size vs number of log lines factor table}

\showgraph{build/graph-input-size-vs-parsing-time}{Parsing time, file size,
and number of log lines}{parsing time vs file size vs number of log lines
graph}

\showgraph{build/graph-input-size-vs-parsing-time-ratio}{Ratio of file size
and number of log lines to parsing time}{parsing time vs file size vs
number of log lines factor}

\subsection{Explaining the peaks in log file size}

@@

\Graphref{parsing time vs file size vs number of log lines graph} shows log
file size varying between 30--100MB, except for log files 22 (200MB) and
62--68 (400--900MB).  These large log files were caused by a user
unintentionally creating a mail forwarding loop, resulting in a huge number
of mail deliveries and up to an order of magnitude more log lines than is
usual.  The user specified his mail forwarding as: \newline{}
\tab{}\texttt{$\backslash$username, username@domain} \newline{} This
instructs Postfix to deliver the mail to the local user, and also forward a
copy to the remote address; this is generally not a problem except that the
remote address in this instance is the address the mail was originally sent
to, creating an infinite loop.  To prevent this happening Postfix examines
the Delivered-To header in the mail, and if the mail has already been
delivered to the current address it is bounced back to the sender with the
error message \newline{} \tab{} \texttt{mail forwarding loop for
username@domain}\newline{}  Ordinarily this works well, but unfortunately
in this instance the user noticed they had not received any mail in a while
and opted to send a test mail to themselves, causing a loop not caught by
Postfix:

XXX REWRITE THIS SECTION SO THAT IT IS CORRECT\@: POINT 4 SAYS THAT THE
ENVELOPE SENDER CHANGES, IS THAT CORRECT\@?

@@

\subsection{Rule ordering for efficiency}

@@

Differences in rule ordering have less effect on parsing time when parsing
log files 22 and 62--68, due to the different distribution of log lines in
those log files (see \sectionref{Explaining the peaks in log file size}).
When a mail loop occurs the vast majority of log lines are from Postfix
components that have few rules associated with them, whereas in general the
most log lines are from Postfix components that have a high number of
associated rules; this results in the average number of rules required to
parse a log line from log files 22 or 62--68 being considerably lower than
usual.  The ratio of number of log lines versus parsing time in
\graphref{parsing time vs file size vs number of log lines factor} shows a
noticeable improvement at log files 22 and 62--68, where a higher ratio
means more log lines parsed per unit of time; similarly \graphref{Parsing
time relative to shuffled ordering graph} shows a much smaller difference
between the orderings at log files 22 and 62--68.

\subsection{Scalability as the number of rules rises}

@@

\newpage{} % XXX
\subsection{Caching compiled regexs}

@@

Two large dips can be seen in \graphref{Increase in parsing time when not
caching compiled regexes graph} at log files 22 and 62--68, corresponding
to the spikes in log file size in \graphref{parsing time vs file size vs
number of log lines graph}.  The distribution of log lines across rules
when there is a mail loop is much different to the norm, and the average
number of rules consulted per log line is much lower: this results in far
fewer regex compilations per line when there is a mail loop, and a
corresponding decrease in the average parsing time for a single log line.
The outcome is that caching compiled regexes is proportionally less
important when the log file contents were created by a mail loop.  The
increases in parsing time when not caching compiled regexes for different
groups of log files are summarised in \tableref{Increase in parsing time
when not caching compiled regexes table}.

@@

@@

\section{Coverage}

@@

\subsection{Log lines covered}

@@

Full coverage of log lines is easy to achieve yet hard to maintain.  It is
easy to achieve full coverage for a limited set of log files (at the time
of writing the parser is tested with \numberOFrules{} rules, fully parsing
\numberOFlogFILESall{} contiguous log files from Postfix 2.2 through to
Postfix 2.5), and new rules are easy to add.  Maintaining full coverage is
hard because other servers have different restrictions with custom
messages, \gls{DNSBL} messages change over time, major releases of Postfix
change warning messages (usually adding more information), etc.,\ so over
time the log lines drift and change.  \Graphref{rule hits graph} shows the
number of hits for each rule over all \numberOFlogFILES{} log files.  The
number of hits per rule is quite unevenly spread, resembling a Power Law
distribution --- it is obvious that a small number of rules match the vast
majority of the log lines, and more than half the rules match fewer than
100 times.

@@

\subsection{Mails covered}

@@

Whereas verifying by inspection that the parser correctly deals with all
60,721,709 log lines in the test log files is infeasible, verifying a
subset of those log files is a tractable, if extremely time consuming,
task.  A sample of log lines was obtained by randomly selecting a log file:

@@

The first 6000 log lines of this log file (roughly 0.01\% of the total
number of log lines used in testing) was extracted:

@@

It is important that the log lines used are contiguous so that all log
entries are present for as many of the connections as possible.  This log
segment was parsed with all debugging options enabled, resulting in 167,448
lines of output.\footnote{A mean of 27.908 lines of output per line of
input; each connection has 30 debugging lines, plus 21 debugging lines per
result.  Connections which have been cloned will have the cloned connection
in their debugging output, plus another 33 debugging lines.  Those numbers
are approximate, and may vary $\pm{}$ 2.  There is a linear relationship
between the number of log lines and debugging lines: $33(connections) +
30(accepted~~mails) + 21(results)$.  This formula is an approximation only,
and has not been rigorously verified.}  All 167,448 lines were examined in
conjunction with the input log file and a dump of the resulting database,
verifying that for each of the log lines the parser used the correct rule
and executed the correct action, which in turn produced the correct result
and inserted the correct data in the database.  The log file segment
produced 4 warnings, 10 mails remaining in the state tables, and 1625
connections correctly entered in the database.

@@

\section{Summary}

### 2009-01-16 doc/logparser.tex
Cleanups and improvements to the review of other parsers.
Proofread and corrected the introduction and LMA section of the state of the art review.
Further corrections and improvements after proofreading.
Further corrections after proofreading.
Further corrections and improvements from proofreading.
Replaced quotation environments around blurbs, they're quite obviously blurbs now.  Other improvements.
Lots of minor improvements and cleanups to the State of the Art Review,     including running diction on it. Checked that I've written 'log files' and 'log lines' everywhere, instead of     'files

% $Id$
\documentclass[a4paper,12pt,draft]{report}

@@

@@

% The name of the program, so I only have to change it in one place.
\newcommand{\parsername}{\gls{PLP}}
% XXX SHOULD THIS BE \glspl{PLP} or \gls{PLP}'s ??
\newcommand{\parsernames}{\glspl{PLP}}

@@

XXX WEKA FOR DECISION TREE DISPLAY\@.

@@

XXX CHECK YEARS FOR BIBLIOGRAPHY ENTRIES\@.

@@

@@

>>>>

% $Id$
% XXX WOULD A SMALLER FONT BE BETTER?
\documentclass[a4paper,12pt,draft]{report}
\pagestyle{headings}

@@

\newcommand{\parserblurb}[3]{%
    \begin{quotation}
        \noindent{}\textit{#1\/}

        \noindent{}\url{#2}\newline
        Last checked #3.
    \end{quotation}
}

@@

% The name of the program, so I only have to change it in one place.
\newcommand{\parsername}{\gls{PLP}}
\newcommand{\parsernames}{\gls{PLP}'s}

@@

@@

XXX CHECK YEARS FOR BIBLIOGRAPHY ENTRIES\@.  WHAT DO I MEAN BY THIS\@?

@@

XXX MAYBE DO SOMETHING ABOUT underfull hboxes IN BIBLIOGRAPHY WHEN URLS
WRAP\@?

@@

XXX EXTRACT DELAY AND DEFER WHEN AUTOMATIC EXTRACTION HAS BEEN FINISHED,
THEN UPDATE FIELDS IN BOTH DATABASE AND REVIEW OF OTHER PARSERS\@.

XXX POSSIBLY PUBLISH THE STATE OF THE ART REVIEW IN A JOURNAL\@?
